---
title: "Opening_NetCDF_daily_and_cumulative"
author: "Elisa"
date: "2024-08-28"
output: word_document
---

Installing or loading packages
```{r}
rm(list=ls()) # cleans the workspace
#install.packages('RNetCDF')
library(RNetCDF)
```

#Daily NetCDF file
```{r}
data <- open.nc("C:\\Users\\uqegirol\\Desktop\\Secondment\\Accelnet project\\NetCDF_files\\Ridgefield_L6_20160101_20240202_Daily.nc", header=TRUE, fill = TRUE) #  imports the file with the data
time <- var.get.nc(data, "time")
precipitation <- var.get.nc(data, "Precip")
soil_water_content <- var.get.nc(data, "Sws")
soil_temperature <- var.get.nc(data, "Ts")
air_temperature <- var.get.nc (data, "Ta")
```

##Converting the date
```{r}
library(lubridate)
origin_date <- as.Date("1800-01-01") # Define the origin date
days_since_1800 <- time
whole_days <- floor(days_since_1800) # Extract the whole number of days
fractional_day <- days_since_1800 - whole_days # Extract the fractional part of days
date_converted <- origin_date + days(whole_days) # Convert the whole number of days to a Date object
fractional_seconds <- fractional_day * 86400 # multiplies the fractional day for the number of seconds in a day. Returns the number of seconds
posix_convertedD <- as.POSIXct(date_converted) + fractional_seconds # Create a POSIXct object with the date and fractional seconds
Tz <- att.get.nc(data, "NC_GLOBAL", "time_zone") # extract the time zone
datetime_converted_tz <- as.POSIXct(posix_convertedD, tz = Tz) # Specify the time zone when creating the POSIXct object. To create a csv file this is commented out because we cannot retain the time zone information in a csv
library(dplyr)
time_zone <- rep(Tz, length(posix_convertedD))
```

##Saving to a dataframe
```{r}
dfD <- data.frame(Time = datetime_converted_tz, Time_zone = time_zone, Precipitation = precipitation, Soil_water_content = soil_water_content, Soil_temperature = soil_temperature, Air_temperature = air_temperature)

# Convert POSIXct to character
dfD$Time <- format(dfD$Time, format = "%Y-%m-%d %H:%M:%S")
```

##Same example plots to check the data
```{r}
plot(posix_convertedD, dfD$Air_temperature)
plot(posix_convertedD, dfD$Precipitation)
```

#Monthly NetCDF file
```{r}
data <- open.nc("C:\\Users\\uqegirol\\Desktop\\Secondment\\Accelnet project\\NetCDF_files\\Ridgefield_L6_20160101_20240202_Monthly.nc", header=TRUE, fill = TRUE) #  imports the file with the data
time <- var.get.nc(data, "time")
precipitation <- var.get.nc(data, "Precip")
soil_water_content <- var.get.nc(data, "Sws")
soil_temperature <- var.get.nc(data, "Ts")
air_temperature <- var.get.nc (data, "Ta")
```

##Converting the date
```{r}
library(lubridate)
origin_date <- as.Date("1800-01-01") # Define the origin date
days_since_1800 <- time
whole_days <- floor(days_since_1800) # Extract the whole number of days
fractional_day <- days_since_1800 - whole_days # Extract the fractional part of days
date_converted <- origin_date + days(whole_days) # Convert the whole number of days to a Date object
fractional_seconds <- fractional_day * 86400 # multiplies the fractional day for the number of seconds in a day. Returns the number of seconds
posix_convertedM <- as.POSIXct(date_converted) + fractional_seconds # Create a POSIXct object with the date and fractional seconds
Tz <- att.get.nc(data, "NC_GLOBAL", "time_zone") # extract the time zone
datetime_converted_tz <- as.POSIXct(posix_convertedM, tz = Tz) # Specify the time zone when creating the POSIXct object. To create a csv file this is commented out because we cannot retain the time zone information in a csv
library(dplyr)
time_zone <- rep(Tz, length(posix_convertedM))
```

##Saving to a dataframe
```{r}
dfM <- data.frame(Time = datetime_converted_tz, Time_zone = time_zone, Precipitation = precipitation, Soil_water_content = soil_water_content, Soil_temperature = soil_temperature, Air_temperature = air_temperature)

# Convert POSIXct to character
dfM$Time <- format(dfM$Time, format = "%Y-%m-%d %H:%M:%S")
```

##Same example plots to check the data
```{r}
plot(posix_convertedM, dfM$Air_temperature)
plot(posix_convertedM, dfM$Precipitation)
```

#Annual NetCDF file
```{r}
data <- open.nc("C:\\Users\\uqegirol\\Desktop\\Secondment\\Accelnet project\\NetCDF_files\\Ridgefield_L6_20160101_20240202_Annual.nc", header=TRUE, fill = TRUE) #  imports the file with the data
time <- var.get.nc(data, "time")
precipitation <- var.get.nc(data, "Precip")
soil_water_content <- var.get.nc(data, "Sws")
soil_temperature <- var.get.nc(data, "Ts")
air_temperature <- var.get.nc (data, "Ta")
```

##Converting the date
```{r}
library(lubridate)
origin_date <- as.Date("1800-01-01") # Define the origin date
days_since_1800 <- time
whole_days <- floor(days_since_1800) # Extract the whole number of days
fractional_day <- days_since_1800 - whole_days # Extract the fractional part of days
date_converted <- origin_date + days(whole_days) # Convert the whole number of days to a Date object
fractional_seconds <- fractional_day * 86400 # multiplies the fractional day for the number of seconds in a day. Returns the number of seconds
posix_convertedA <- as.POSIXct(date_converted) + fractional_seconds # Create a POSIXct object with the date and fractional seconds
Tz <- att.get.nc(data, "NC_GLOBAL", "time_zone") # extract the time zone
datetime_converted_tz <- as.POSIXct(posix_convertedA, tz = Tz) # Specify the time zone when creating the POSIXct object. To create a csv file this is commented out because we cannot retain the time zone information in a csv
library(dplyr)
time_zone <- rep(Tz, length(posix_convertedA))
```

##Saving to a dataframe
```{r}
dfA <- data.frame(Time = datetime_converted_tz, Time_zone = time_zone, Precipitation = precipitation, Soil_water_content = soil_water_content, Soil_temperature = soil_temperature, Air_temperature = air_temperature)

# Convert POSIXct to character
dfA$Time <- format(dfA$Time, format = "%Y-%m-%d %H:%M:%S")
```

##Same example plots to check the data
```{r}
plot(posix_convertedA, dfA$Air_temperature)
plot(posix_convertedA, dfA$Precipitation)
```
#Cumulative NetCDF file
```{r}
data <- open.nc("C:\\Users\\uqegirol\\Desktop\\Secondment\\Accelnet project\\NetCDF_files\\Ridgefield_L6_20160101_20240202_Cumulative.nc", header=TRUE, fill = TRUE) #  imports the file with the data
time <- var.get.nc(data, "time")
precipitation <- var.get.nc(data, "Precip")
```

##Converting the date
```{r}
library(lubridate)
origin_date <- as.Date("1800-01-01") # Define the origin date
days_since_1800 <- time
whole_days <- floor(days_since_1800) # Extract the whole number of days
fractional_day <- days_since_1800 - whole_days # Extract the fractional part of days
date_converted <- origin_date + days(whole_days) # Convert the whole number of days to a Date object
fractional_seconds <- fractional_day * 86400 # multiplies the fractional day for the number of seconds in a day. Returns the number of seconds
posix_convertedC <- as.POSIXct(date_converted) + fractional_seconds # Create a POSIXct object with the date and fractional seconds
Tz <- att.get.nc(data, "NC_GLOBAL", "time_zone") # extract the time zone
datetime_converted_tz <- as.POSIXct(posix_convertedC, tz = Tz) # Specify the time zone when creating the POSIXct object. To create a csv file this is commented out because we cannot retain the time zone information in a csv
library(dplyr)
time_zone <- rep(Tz, length(posix_convertedC))
```

##Saving to a dataframe
```{r}
dfC <- data.frame(Time = datetime_converted_tz, Time_zone = time_zone, Precipitation = precipitation)

# Convert POSIXct to character
dfC$Time <- format(dfC$Time, format = "%Y-%m-%d %H:%M:%S")
```

##Same example plots to check the data
```{r}
plot(posix_convertedC, dfC$Precipitation)
```
